---
title: 'HEALTH mom data analysis'
author: "Alejandra Muñoz-Rivera"
date: '4/9/23'
---
#Open librarys 
```{r}
pacman::p_load(VIM, mice, lattice, ggplot2, tidyverse, foreign, table1, haven, labelled, boot, tableone, readr, nnet, MASS, funModeling, brant, broom, finalfit, dplyr, knitr, kableExtra, gridExtra, boot, multcomp, lmtest, odds.n.ends, blorr, car, jtools, broom, openxlsx, brant)
```
 
#import datasets 
```{r}
#import demographic baseline parent dataset 
par_bl <-read_sas("/Users/alemunoz/Desktop/BrownSchool/Fall22/Capstone/Data/par_bl_12sep2022.sas7bdat")

#import PSI baseline parent dataset 
par_psi <-read_sas("/Users/alemunoz/Desktop/BrownSchool/Fall22/Capstone/Data/par_psi_12sep2022.sas7bdat")


```

#Clean data Mom BL data 

```{r}

var_label(par_bl) <- NULL #drop labels 

#rename funny labelled variables
par_bl<-rename(par_bl, c('ethnicity'='_ethnicity', 'race'='_race'))
par_bl<-rename(par_bl, c('visit' = '_visit'))
par_bl<-rename(par_bl, c('fi' = '_fi'))
par_bl<-rename(par_bl, c('pp_meals' = '_pp_meals'))
par_bl<-rename(par_bl, c('aa_food' = '_aa_food'))

#recode variables as factor 
Moml$ethnicity <-as.factor(Mom$ethnicity)
Mom$race <-as.factor(Mom$race)


```


```{r}
#Create smaller dataset w/ variables of interest from demographic dataset 
Mom <-par_bl %>% 
  dplyr::select(study_id,dem1, ethnicity, race, dem5, dem6, dem7, fi) #

#recode variables as factor 
Mom$ethnicity <-as.factor(Mom$ethnicity)
Mom$race <-as.factor(Mom$race)
class(Mom$ethnicity)
class(Mom$race)

#recode all variables 
Mom <- Mom %>%
    mutate(raceth = case_when (ethnicity == 1 ~ 0, #Hispanic/Latino
                             ethnicity %in% c(2,99) & race == 0 ~ 1, #American Indian/Alaskan Native
                             ethnicity %in% c(2,99) & race %in% c(1,2) ~ 2, #Asian/Pacific Islander
                             ethnicity %in% c(2,99) & race == 3 ~ 3, #Black or African American
                             ethnicity %in% c(2,99) & race == 4 ~ 4, #White
                             ethnicity %in% c(2,99) & race == 5 ~ 5)) %>% #More than one race 
      mutate(raceth = recode_factor(.x = raceth, 
                                  `0` = "Hispanic/Latino",
                                  `1` = "American Indian/Alaskan Native", 
                                  `2` = "Asian/Pacific Islander",
                                  `3` = "Black or African American",
                                  `4` = "White",
                                  `5` = "More than one race ")) %>%
      mutate(marital = recode_factor(.x = dem6, 
                                  `1` = "Single",
                                  `2` = "Married or living w/ partner", 
                                  `3` = "Seperated/Divorced",
                                  `4` = "Seperated/Divorced",
                                  `5` = "Widow",
                                  `6` = "Married or living w/ partner"))%>%
      mutate(income = recode_factor(.x = dem7, 
                                    `1` = "Under $10,000",
                                    `2` = "$10,000-$29,999",
                                    `3` = "$10,000-$29,999",
                                    `4` = "$30,000-$49,999",
                                    `5` = "$30,000-$49,999", 
                                    `6` = "$50,000-$74,999", 
                                    `7` = "$75,000-$99,999",
                                    `8` = "Over $100,00")) %>%
        mutate(fi = recode_factor(.x = fi, 
                                  `1` = "Food Insecure", 
                                  `0` = "Food Secure")) %>%
        mutate(edu = recode_factor(.x = dem5, 
                                  `1` = "Some high school or less", 
                                  `2` = "Some high school or less",
                                  `3` = "High school graduate or GED",
                                  `4` = "Some college",
                                  `5` = "Technical or vocational school",
                                  `6` = "College or university graduate", 
                                  `7` = "Graduate or professional school")) 



```

#Clean PSI data set 
```{r}
var_label(par_psi) <- NULL #drop labels 
#rename funky variable names 
par_psi<-rename(par_psi, c('visit'='_visit', 'trt' = '_trt')) 
par_psi<-rename(par_psi, c('date' = '_max_date'))

#create smaller dataset for PSI data
psi <- par_psi %>% dplyr::select(study_id, psi5, psi6, psi7, psi8, psi9, psi10, )

#recode PSI variables of interest
psi <- psi %>%
  mutate(snacks= recode_factor(.x = psi5, 
                                  `3` = "Stayed the same",
                                  `4` = "Stayed the same",   
                                  `1` = "Increased",
                                  `2` = "Decreased"))%>%
  mutate(veggies= recode_factor(.x = psi6, 
                                  `3` = "Stayed the same",
                                  `4` = "Stayed the same",
                                  `1` = "Increased",
                                  `2` = "Decreased"))%>%
  mutate(eating= recode_factor(.x = psi7, 
                                  `3` = "Stayed the same",
                                  `4` = "Stayed the same",
                                  `1` = "Less Healthy",
                                  `2` = "More Healthy",))%>%
  mutate(weight= recode_factor(.x = psi8, 
                                  `3` = "Stayed the same",
                                  `4` = "Stayed the same",
                                  `1` = "Gained weight",
                                  `2` = "Lost weight"))%>%
  mutate(incomech= recode_factor(.x = psi9, 
                                  `1` = "No change",
                                  `2` = "Small change",
                                  `3` = "Moderate change",
                                  `4` = "Severe change"))%>%
  mutate(faccess= recode_factor(.x = psi10, 
                                  `1` = "No change",
                                  `2` = "Small change",
                                  `3` = "Moderate change",
                                  `4` = "Severe change"))

psi <- drop_na(psi)

```

```{r}
#Merge bl and psi by study ID 
total <- merge(Mom, psi, by=c("study_id"), all.x = TRUE, all.y = TRUE)
total <- total %>% dplyr::select(-c(dem5, dem6, dem7, psi5, psi6, psi7, psi8, psi9, psi10, faccess, Age, ethnicity, race, raceth))
total <-drop_na(total)

#missing data patterns 
md.pattern(total, rotate.names=TRUE)
md.pairs(total)

#impute data 
total_in1<-droplevels(total)
total_in2<-mice(total, m=5, maxit=5, seed=219) 
total_in2
print(total_in2)
```


# Table1 of descriptives 
```{r}
## Add labels## 
label(total$fi)<-"Food Security Status"
label(total$ethnicity)<-"Ethnicity"
label(total$race)<-"Race"
label(total$marital)<-"Marital Status"
label(total$incomech)<-"Change in income"
label(total$income) <- "Annual Household Income"
label(total$edu) <- "Education level"
label(total$raceth) <- "Race/Ethnicity"

#descriptives table 
table1(~Age + raceth + marital + income + edu + fi, Mom)

```


#Bivariate table (needs to be edited)
```{r echo=FALSE, warning=FALSE,include=FALSE, message=FALSE }
cor.test(total$psi5, total$psi7, method=c("kendall"))
cor.test(total$psi6, total$psi7, method=c("kendall"))
```
Low correlation between psi5 and psi7 and between psi6 and psi7. Therefore, I will run two seperate multinomial regression with psi5 and psi6 as outcomes, respectively. Food security status will be the predictor. Since there is a low response rate for loss of income I will run analysis with compelte cases only and inputed data. 

#Models 
```{r}
#model with incidence of snacks with complete data only
model1 <- polr(snacks ~ fi + incomech + edu, Hess=TRUE, data=total)
# odds ratios and 95%CIs
data <- tidy(model1, conf.int=TRUE, exponentiate = TRUE)


#model with incidence of snacks with inputed data
model1imp <- with(total_in2, multinom(snacks ~ fi + incomech + edu))
#odds ratios and 95CIs - calculated log odds and exponentiate 
Table2 <- summary(pool(model1imp),  conf.int = TRUE, exponentiate = TRUE)
Table2

```
Food security status only incidence of snacks with inputed data when comparing decreasing snack incidence vs staying the same. 


```{r}
#model with incidence of fruits and veggies with complete data only
model2 <- polr(veggies ~ fi + incomech + edu, Hess=TRUE, data=total1)
# odds ratios and 95%CIs
tidy(model4, conf.int=TRUE, exponentiate = TRUE, p.values=TRUE)

#model with fruits and veggies with inputed data
model2imp <- with(total_in2, multinom(veggies ~ fi + incomech + edu))
#odds ratios and 95CIs - calculated log odds and exponentiate 
Table3 <- summary(pool(model2imp),  conf.int = TRUE, exponentiate = TRUE)
Table3
```
Food security does not predict incidence of fruit and veggie consumption with either dataset. 

```{r}
#model with incidence of eating habits with complete data only
model3 <- polr(eating ~ fi + incomech + edu, Hess=TRUE, data=total1)
#odds ratios and 95%CIs
tidy(model3, conf.int=TRUE, exponentiate = TRUE, p.values=TRUE)

#model with incidence of eating habits with inputed data
model3imp <- with(total_in2, multinom(eating ~ fi + incomech + edu))
#odds ratios and 95CIs - calculated log odds and exponentiate 
Table4 <- summary(pool(model3imp),  conf.int = TRUE, exponentiate = TRUE)
Table4

```
Food security does not predict precieved eating habits with either data set. 


```{r}
#model with precieved weight change with complete data only
model4 <- multinom(weight ~ fi + incomech + edu, data=total)
# odds ratios and 95%CIs
tidy(model4, conf.int=TRUE, exponentiate = TRUE, p.values=TRUE)

#model with precieved weight change with inputed data
model4imp <- with(total_in2, multinom(weight ~ fi + incomech + edu, Hess=TRUE))
#odds ratios and 95CIs - calculated log odds and exponentiate 
Table5 <- summary(pool(model4imp),  conf.int = TRUE, exponentiate = TRUE)
Table5

```
Food security status does not predict incidence of snacks with either data set. 



